noam_final_analyses

### 2024-09-06 noam_final_analyses.Rmd
added model selection for macroalgae, but I can't get stat summaries for forereef data even though there are no warnings in the model or the dredge.

alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))

@@

## FRINGING - exclude wave power
fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors

fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) # universal model has convergence issues
fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) # most models have convergence issues
model.sel(fringing_all_richness_dredge) # top model just includes year
fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass)
summary(fringing_all_richness_top_mod)
hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine

fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
model.sel(fringing_all_shannon_dredge) # tie for the top model - best includes year and coral cover or just nitrogen
fringing_all_shannon_top_mod1 <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_top_mod2 <- glmmTMB(log(shannon + 1) ~ mean_n + mean_temp_c + (1|site), data = fringing_all_site, na.action = na.pass)
hist(residuals(fringing_all_shannon_top_mod1)) # slightly bimodal
hist(residuals(fringing_all_shannon_top_mod2)) # also uniform - neither are great
plot(residuals(fringing_all_shannon_top_mod1) ~ fitted(fringing_all_shannon_top_mod1)) # slight patterning
plot(residuals(fringing_all_shannon_top_mod2) ~ fitted(fringing_all_shannon_top_mod2)) # similar here

## Backreef - includes all 
backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors

backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has convergence issues
backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
model.sel(backreef_all_richness_dredge) # top model includes year and coral 
backreef_all_richness_top_mod <- glmmTMB(richness ~ year + coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass)
summary(backreef_all_richness_top_mod)
hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
model.sel(backreef_all_shannon_dredge) #  best includes just coral cover 
backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  coral_cover_hard + (1|site), data = backreef_all_site, na.action = na.pass)
hist(residuals(backreef_all_shannon_top_mod)) # pretty good
plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine

@@

outer10_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence issues
outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
model.sel(outer10_all_richness_dredge) # top model includes just intercept... next best is coral 
outer10_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass)
summary(outer10_all_richness_top_mod)
hist(residuals(outer10_all_richness_top_mod)) #  not bad
plot(residuals(outer10_all_richness_top_mod) ~ fitted(outer10_all_richness_top_mod)) # looks fine

outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) # some convergence issues
model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness - first time there's been a decent gap in AICc
outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
hist(residuals(outer10_all_shannon_top_mod)) # pretty good
plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine

@@

outer17_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
model.sel(outer17_all_richness_dredge) # top model includes just intercept... next best is coral 
outer17_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass)
summary(outer17_all_richness_top_mod)
hist(residuals(outer17_all_richness_top_mod)) #  not bad
plot(residuals(outer17_all_richness_top_mod) ~ fitted(outer17_all_richness_top_mod)) # looks fine

outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
model.sel(outer17_all_shannon_dredge) #  best includes just year
outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
hist(residuals(outer17_all_shannon_top_mod)) # pretty good
plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
```

>>>>

alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))
alpha_diversity_site_macro$richness_area <- alpha_diversity_site_macro$richness/(alpha_diversity_site_macro$hard_cover/100)

@@

## FRINGING - exclude wave power
fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors
# 
# fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + mean_n + mean_temp_c +  herbivore_evenness +
# offset(log(hard_cover)) #+ (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) 
# fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) 
# model.sel(fringing_all_richness_dredge) # top model just includes year
# fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = # fringing_all_site, na.action = na.pass)
# summary(fringing_all_richness_top_mod)
# hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
# plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine

fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
model.sel(fringing_all_shannon_dredge) # tie for the top model just nitrogen
fringing_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ mean_n + (1|site), data = fringing_all_site, na.action = na.pass)

qqnorm(residuals(fringing_all_shannon_top_mod))# not bad
qqline(residuals(fringing_all_shannon_top_mod))
plot(residuals(fringing_all_shannon_top_mod) ~ fitted(fringing_all_shannon_top_mod)) # slight patterning

## Backreef - includes all 
# backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
# 
# backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has # convergence issues
# backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
# model.sel(backreef_all_richness_dredge) # top model includes year and coral 
# backreef_all_richness_top_mod <- glmmTMB(richness ~ year  + offset(log(hard_cover)) + (1|site), family = poisson, data = # backreef_all_site, na.action = na.pass)
# summary(backreef_all_richness_top_mod)
# hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
# plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
model.sel(backreef_all_shannon_dredge) #  best includes just year, next best also includes herbivore evenness
backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = backreef_all_site, na.action = na.pass)
qqnorm(residuals(backreef_all_shannon_top_mod)) #  a little wonky on the tail but not bad
plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine
summary(backreef_all_shannon_top_mod)

@@

# outer10_all_richness_mod_universal <- glmmTMB(richness ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence # issues
# outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
# model.sel(outer10_all_richness_dredge)

# 
outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) 
model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness 
outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
qqnorm(residuals(outer10_all_shannon_top_mod)) # pretty good
plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine
summary(outer10_all_shannon_top_mod)

@@

# outer17_all_richness_mod_universal <- glmmTMB(richness ~ year  + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + # (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
# outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
# model.sel(outer17_all_richness_dredge) # top model includes just intercept...

outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
model.sel(outer17_all_shannon_dredge) #  best includes just year. next best also includes temperature and is almost equivalent AICc
outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
qqnorm(residuals(outer17_all_shannon_top_mod)) # pretty good
plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
summary(outer17_all_shannon_top_mod)

```{r stepwise models site level macroalgae}

## FRINGING - exclude wave power
fringing_macro_site <- na.omit(subset(alpha_diversity_site_macro, habitat == "fringing")) # only use rows that contain all factors

fringing_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_macro_site, na.action = na.pass)
fringing_macro_shannon_dredge <- dredge(fringing_macro_shannon_mod_universal) 
model.sel(fringing_macro_shannon_dredge) # top model just year, next best includes temp
fringing_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year + (1|site), data = fringing_macro_site, na.action = na.pass)
qqnorm(residuals(fringing_macro_shannon_top_mod))# not bad
plot(residuals(fringing_macro_shannon_top_mod) ~ fitted(fringing_macro_shannon_top_mod)) # slight negative trend

## Backreef - includes all 
backreef_macro_site <- na.omit(subset(alpha_diversity_site_macro, habitat == "backreef")) # only use rows that contain all factors

backreef_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_macro_site, na.action = na.pass)
backreef_macro_shannon_dredge <- dredge(backreef_macro_shannon_mod_universal) 
model.sel(backreef_macro_shannon_dredge) #  best includes just wave power but estimate is tiny
backreef_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  wave_power_adcp + (1|site), data = backreef_macro_site, na.action = na.pass)
qqnorm(residuals(backreef_macro_shannon_top_mod)) #  a little wonky but not bad
plot(residuals(backreef_macro_shannon_top_mod) ~ fitted(backreef_macro_shannon_top_mod)) # looks fine
summary(backreef_macro_shannon_top_mod)
Anova(backreef_macro_shannon_top_mod) # lol X2 = 2.56, P = 0.1

## outer10 - no nitrogen 
outer10_macro_site <- subset(alpha_diversity_site_macro, habitat == "outer_10") # only use rows that contain all factors
# remove nitrogent columns
outer10_macro_site<- outer10_macro_site[,-which(names(outer10_macro_site) %in% c("mean_n", "n_cv"))]
# remove other na rows
outer10_macro_site <- na.omit(outer10_macro_site)

outer10_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_macro_site, na.action = na.pass)
outer10_macro_shannon_dredge <- dredge(outer10_macro_shannon_mod_universal) 
model.sel(outer10_macro_shannon_dredge) #  best includes wave power and year 
outer10_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year  + (1|site), data = outer10_macro_site, na.action = na.pass)
qqnorm(residuals(outer10_macro_shannon_top_mod)) # pretty good, tails maybe off
plot(residuals(outer10_macro_shannon_top_mod) ~ fitted(outer10_macro_shannon_top_mod)) # looks fine
summary(outer10_macro_shannon_top_mod) # NaN?

## outer17 - no nitrogen or herbivores
outer17_macro_site<- subset(alpha_diversity_site_macro, habitat == "outer_17") # only use rows that contain all factors
# remove nitrogent columns
outer17_macro_site<- outer17_macro_site[,-which(names(outer17_macro_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
# remove other na rows
outer17_macro_site <- na.omit(outer17_macro_site)

outer17_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_macro_site, na.action = na.pass)
outer17_macro_shannon_dredge <- dredge(outer17_macro_shannon_mod_universal) # some convergence issues
model.sel(outer17_macro_shannon_dredge) #  best includes all terms!
qqnorm(residuals(outer17_macro_shannon_mod_universal)) # a little hump shaped
plot(residuals(outer17_macro_shannon_mod_universal) ~ fitted(outer17_macro_shannon_mod_universal)) # looks fine
summary(outer17_macro_shannon_mod_universal) # also NaNs...
